Benchmarking on Rigetti architecture

### 2022-06-23 txt-softwarex/SoftwareX_article_template.tex
add theorems to the section about example

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we will compare theoretical probability of correct 
discrimination between these measurements and the realization on real devices.
The theoretical probability can be calculated using Holevo-Helstrom theorem. 
This probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme on real devices in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

\begin{proposition}
 Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
 $\Phi_U$ and $\Phi_\Id$ be two unitary channels. The following equation holds 
	\begin{equation}
	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond
	\end{equation}
\end{proposition}
\begin{proof} We know that
$
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
$
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. For $U = H_2 \left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H_2^\dagger$ we obtain that $\nu = \frac{|1 + e^{i \phi} | }{2}$. 
	The condition $ 	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond $ is equivalent to prove that
	\begin{equation}
	\max_{E \in \diaguni_2 } \nu_E = \nu = \frac{|1 + e^{i \phi} | }{2},
	\end{equation}
	where $\nu_E = \min_{x \in W(U^\dagger E)} |x|. $ We can prove that
	\begin{equation}
	\min_{\ket{x} \in \mathbb{C}^2  \proj{x} = 1} |\bra{x}U^\dagger\ket{x}| = \min_{\rho \in \Omega_2} |\tr(U^\dagger\rho)|. 
	\end{equation}
For that, to calculate $\nu_E$ we obtain \begin{equation}
	\max_{E \in \diaguni_2 } \nu_E  = \max_{E \in \diaguni_2 }  \min_{\rho \in \Omega_2} \left| \tr \rho U E  \right|
\end{equation}
For that, our task is reduce to show that
	\begin{equation}
	\forall E \in \diaguni_2 \,\, | \tr \rho U E | \le \nu. 
	\end{equation}
	Let us define $E = \left(\begin{array}{cc}E_0&0\\0&E_1\end{array}\right)  $ and let us take $\rho = \left(\begin{array}{cc}\frac{1}{2}&0\\0&\frac{1}{2}\end{array}\right) $. 
	From spectral theorem, let us note $U$ as
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvector $\lambda_1 = e^{\mathbf{i} \phi}$ the eigenvector is of the form $\ket{x_1} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right]$, whereas for  $\lambda_2 = 1$ we have $\ket{x_2} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]$.

@@

\begin{remark}
	The probability of correct discrimination of von Neumann measurements 
	$\PP_U$ and $\PP_{\Id}$ for $U = H_2 \diag(1, e^{i \phi}) H_2^\dagger$, 
	where $\phi \in [0, 2\pi)$ is given by
	\begin{equation}
	p \le \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
	\end{equation}
\end{remark}

\subsection{Realization}\label{sec:example_realization}

@@

@@

\subsubsection{Final measurement}\label{sec_example_final_measurement}
\begin{scheme}
Let us consider the problem of discrimination of von Neumann measurements 
$\PP_U$
and $\PP_{\Id}$ for $U = H_2 \diag(1, e^{i \phi}) H_2^\dagger$, 
	where $\phi \in [0, 2\pi)$. The schematic representation of theoretical 
	setup is shown in Fig.~\ref{fig:theoretical}.

From ~\cite{puchala2018strategies}(Proposition 4) and Lemma~\ref{lemma:rho} we 
have 
	\begin{equation}
	\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
	\rangle \rangle. 
	\end{equation}
	From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
	Let us define \begin{equation}
	X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
	\otimes \Id_2 \right)(\proj{\psi})
	\end{equation}
	where $\ket{\psi}$ is defined in Remark~\ref{remark:discriminator}. From 
	Hahn-Jordan decomposition let \begin{equation}
	X = P - Q
	\end{equation}
	where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
	Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $im(P)$ and $im(Q)$, 
	respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
	\begin{equation}
	\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
	\end{equation}
	and 
	\begin{equation}
	\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
	\end{equation}
	Hence, we define $V_0$ such that
	\begin{equation}
	\begin{split}
	V_0 \ket{x_p} = \ket{0} \\ 
	V_0 \ket{x_q} = \ket{1}
	\end{split}
	\end{equation}
	and $V_1$ such that
	\begin{equation}
	\begin{split}
	V_1 \ket{y_p} = \ket{0} \\ 
	V_1 \ket{y_q} = \ket{1}
	\end{split}
	\end{equation}
	The explicit form of $V_0$ and $V_1$ we can see in \texttt{one-qubit.nb}.
	Finally, we obtain the measurement $\mu$ of the form
	\begin{equation}
	\begin{split}
	 \mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
	 \proj{0} V_1^\dagger  \\ 
	\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{1} V_1^\dagger  
	\end{split}
	\end{equation}
\end{scheme}

@@

>>>>

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we will compare theoretical probability of correct 
discrimination between these measurements and the realization on real devices.
The theoretical probability can be calculated using Holevo-Helstrom theorem and 
Theorem [] from~\cite{puchala2018strategies}. 
This probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme on real devices in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

\begin{theorem}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ for $U = H \diag(1, e^{i \phi}) H^\dagger$,
where $\phi \in [0, 2\pi)$ is given by
\begin{equation}
p = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}

\todo[inline]{change proof of proposition into proof of theorem}
\begin{proposition}
 Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
 $\Phi_U$ and $\Phi_\Id$ be two unitary channels. The following equation holds 
	\begin{equation}
	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond
	\end{equation}
\end{proposition}
\begin{proof} We know that
$
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
$
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. For $U = H_2 \left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H_2^\dagger$ we obtain that $\nu = \frac{|1 + e^{i \phi} | }{2}$. 
	The condition $ 	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond $ is equivalent to prove that
	\begin{equation}
	\max_{E \in \diaguni_2 } \nu_E = \nu = \frac{|1 + e^{i \phi} | }{2},
	\end{equation}
	where $\nu_E = \min_{x \in W(U^\dagger E)} |x|. $ We can prove that
	\begin{equation}
	\min_{\ket{x} \in \mathbb{C}^2  \proj{x} = 1} |\bra{x}U^\dagger\ket{x}| = \min_{\rho \in \Omega_2} |\tr(U^\dagger\rho)|. 
	\end{equation}
For that, to calculate $\nu_E$ we obtain \begin{equation}
	\max_{E \in \diaguni_2 } \nu_E  = \max_{E \in \diaguni_2 }  \min_{\rho \in \Omega_2} \left| \tr \rho U E  \right|
\end{equation}
For that, our task is reduce to show that
	\begin{equation}
	\forall E \in \diaguni_2 \,\, | \tr \rho U E | \le \nu. 
	\end{equation}
	Let us define $E = \left(\begin{array}{cc}E_0&0\\0&E_1\end{array}\right)  $ and let us take $\rho = \left(\begin{array}{cc}\frac{1}{2}&0\\0&\frac{1}{2}\end{array}\right) $. 
	From spectral theorem, let us note $U$ as
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvector $\lambda_1 = e^{\mathbf{i} \phi}$ the eigenvector is of the form $\ket{x_1} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right]$, whereas for  $\lambda_2 = 1$ we have $\ket{x_2} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]$.

@@

\subsection{Realization}\label{sec:example_realization}
Now we focus on the realization of the discrimination protocol. We will 
discriminate between $\PP_U$ and $\PP_\1$. To do so, we need to know exact 
forms of the input state and final measurement. We will indicate the optimal 
input state in subsection~\ref{sec:example_discriminator} and the final 
measurement in subsection~\ref{sec_example_final_measurement}.

@@

\begin{proposition}
Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
and $\PP_\1$, where $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
2\pi)$.  The optimal input state has the form
\begin{equation}
\ket{\psi_{AB}} = \frac{1}{\sqrt{2}} |\Id_2 \rangle \rangle.
\end{equation}
\end{proposition}

\todo[inline]{add formal proof of the proposition}

@@

\subsubsection{Final measurement}\label{sec_example_final_measurement}
In this subsection we will derive exact form of the final measurement in the
problem of discrimination of von Neumann measurements $\PP_U$ and $\PP_{\Id}$
for $U = H \diag(1, e^{i \phi}) H^\dagger$, where $\phi \in [0, 2\pi)$.

\begin{proposition}
The  final measurement has the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger,
\end{split}
\end{equation}
where for each $\phi \in \mathbb{R}$,  the controlled unitaries $V_0$ and $V_1$ 
have the form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right) 
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
\end{proposition}
\todo[inline]{add proof of the proposition}

%The schematic representation of theoretical 
%setup is shown in Fig.~\ref{fig:theoretical}.

From ~\cite{puchala2018strategies}(Proposition 4) and Lemma~\ref{lemma:rho} we 
have 
\begin{equation}
\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
\rangle \rangle. 
\end{equation}
From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
Let us define \begin{equation}
X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
\otimes \Id_2 \right)(\proj{\psi})
\end{equation}
where $\ket{\psi}$ is defined in Remark~\ref{remark:discriminator}. From 
Hahn-Jordan decomposition let \begin{equation}
X = P - Q
\end{equation}
where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $im(P)$ and $im(Q)$, 
respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
\begin{equation}
\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
\end{equation}
and 
\begin{equation}
\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
\end{equation}
Hence, we define $V_0$ such that
\begin{equation}
\begin{split}
V_0 \ket{x_p} = \ket{0} \\ 
V_0 \ket{x_q} = \ket{1}
\end{split}
\end{equation}
and $V_1$ such that
\begin{equation}
\begin{split}
V_1 \ket{y_p} = \ket{0} \\ 
V_1 \ket{y_q} = \ket{1}
\end{split}
\end{equation}
The explicit form of $V_0$ and $V_1$ we can see in \texttt{one-qubit.nb}.
Finally, we obtain the measurement $\mu$ of the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger  
\end{split}
\end{equation}

@@

\subsubsection{Two schemes of practical realization}
